---
title: "Assignment I - CompStat2023"
author: "Aldo Giovanni e Giacomo"
date: "Anuja Saira Abraham 5204982, Alessia Marzotti 5108443, Miriam Mercuri 5207057"
output:
<<<<<<< Updated upstream
  pdf_document: default
  html_document: default
=======
  pdf_document:
    toc: yes
  html_document:
    theme: readable
    toc: yes
    code_folding: show
    df_print: paged
editor_options: 
  markdown: 
    wrap: 72
>>>>>>> Stashed changes
---

# Set up

```{r}

```

# Exeercise 1

## Point a

Compute $P [\max(X_1, X_2) > Y_1]$.

We know that $Z=max(X_1,X_2) \in (1,2,3,4,5,6)$ and that
$P(Z=z)=(2(z-1)+1)/36$. In other word we want to find
$P(Z>y)=1-P(Z \leq y)$.
$P(Z \leq y)=\\P(Z=1 \cap y=1)+\\P(Z=1 \cap y=2)+P(Z=2 \cap y=2)+\\P(Z=1 \cap y=3)+P(Z=2 \cap y=3)+P(Z=3 \cap y=3)+\\P(Z=1 \cap y=4)+P(Z=2 \cap y=4)+P(Z=3 \cap y=4)+P(Z=4 \cap y=4)+\\P(Z=1 \cap y=5)+P(Z=2 \cap y=5)+P(Z=3 \cap y=5)+P(Z=4 \cap y=5)+P(Z=5 \cap y=5)+\\P(Z=1 \cap y=6)+P(Z=2 \cap y=6)+P(Z=3 \cap y=6)+P(Z=4 \cap y=6)+P(Z=5 \cap y=6)+P(Z=6 \cap y=6)$

So we have
$P(Z \leq y)= 6\frac{1}{36}\frac{1}{6}+5\frac{3}{36}\frac{1}{6}+4\frac{5}{36}\frac{1}{6}+3\frac{7}{36}\frac{1}{6}+2\frac{9}{36}\frac{1}{6}+\frac{11}{36}\frac{1}{6}\approx0,421$
and $$P(Z>y)=1 - 0,421=0,579$$

## Point b-c-d

Here there are some code to simulate a generic Risiko! game for
different values of competing units.

```{r}
library(ggplot2)

set.seed(123)
combat_round <- function(att_units,def_units,sim=10000) {
  Results = rep(NA,sim)
  AS<-att_units
  DS<-def_units
  for(i in 1:sim){
    while(def_units>0 & att_units>0){
      Dnum <- sort(sample(1:6, min(def_units,3),replace = TRUE),decreasing = T)
      Anum <- sort(sample(1:6, min(att_units,3),replace = TRUE),decreasing = T)
      for (j in 1:min(length(Dnum),length(Anum))){
        if(Anum[j]>Dnum[j]){
          def_units<-def_units-1
        }
        else{
          att_units<-att_units-1
        }
      }
    }
    Results[i]<- ifelse(att_units>0,1,0)
    att_units<-AS
    def_units<-DS
  }
  return(mean(Results))
}

prob_att_win <- function(){
  res<-sapply(1:10, function(x) {
  sapply(1:10, function(y) {
    combat_round(y,x,1000)
    })
  })
  return(res)
}
# Print the result
#print(result)


att_prob<- data.frame(
  attacker_unit<- rep(1:10,10),
  defender_unit<- rep(1:10,each= 10),
  win_prob<- as.vector(prob_att_win())
)


#go through outer 
ggplot(att_prob,aes(x=defender_unit,y=attacker_unit,fill=win_prob))+
  geom_tile()+
  scale_fill_gradient(low="grey",high="darkgreen")+
  labs(title = "Attacker Win Probability",x="Defender Units",y="Attacker Units")+
  theme_minimal()



print(combat_round(def_units=1,att_units=2,sim=10000))
```

<<<<<<< Updated upstream


=======
>>>>>>> Stashed changes
# Exercise 3

## Point a

The Pareto distribution is defined by a density
$f(x;\gamma)=\gamma x^{-(\gamma +1)}$ over $(1;+\infty)$, with
$\gamma >0$.

It can be generated as the $-\frac{1}{\gamma}$ power of a uniform r.v.

Cumulative distribution function of Pareto distribution
$\int_1^x \gamma z^{-(\gamma +1)} dz= \gamma \int_1^x z^{-1-\gamma}dz=-[x^{-\gamma}-1^{-\gamma}]=-x^{-\gamma}+1=1-(\frac{1}{x})^{\gamma}$

We will use the following theorem: if $X \sim F(x)$ then
$U=F(x) \sim U(0,1)$ $F(X)=1-(\frac{1}{x})^{\gamma}=U$
$(1-U)^{-\frac{1}{\gamma}}=(x^{\gamma})^{-\frac{1}{\gamma}}$
$x=(1-U)^{-\frac{1}{\gamma}}=U^{-\frac{1}{\gamma}}$

## Pont b

Reference: Wikipedia The Pareto distribution is related to the
exponential distribution as follows. If X is Pareto-distributed with
minimum $x_m$ and index $\alpha$, then $Y=log(\frac{X}{x_m})$ is
exponentially distributed with rate parameter $\alpha$.

$Y=log(\frac{X}{x_m}) \sim Exp(\gamma)$ $Y=log(X) \sim Exp(\gamma)$

##Point c

Implement two samplers, one for X and one for Y . Plot the histogram and
the density and comment on the results exploring different values of γ.

```{r}
#sampler for Pareto
xsampler<- function(n, gamma){
  u<- runif(n)
  1/((1-u)^(1/gamma))
}

#sampler for exponential
ysampler<- function(n,gamma){
  x<- xsampler(n,gamma)
  y<- log(x)
  return(y)
}

x1=xsampler(10000,40)
y1=ysampler(10000,40)
hist(x1,breaks = 30, freq=F)
hist(y1,breaks = 30, freq=F)
```

# Exercise 4

## Point a

We write the probability of a standard Normal r.v. X as an integral
thanks to its probability density function.

$P(X>20)=\int_{20}^\infty \frac{1}{\sqrt{2\pi}} e^{-\frac{x^{2}}{2}} dx$

Reference: Wikipedia

The crude Monte Carlo estimation of this quantity is deemed to fail
because the region of integration is so far out in the tails of the
standard normal distribution.

## Point b

Rewrite the integral employing the change of variable $Y = \frac{1}{X}$.

$dy=-\frac{1}{x^{2}} dx\\dx=-x^{2}dy=-\frac{1}{y^{2}}dy\\y=Y_{20}=0,05\\y=Y_\infty=0$

So
$\int_{20}^\infty \frac{1}{\sqrt{2\pi}} e^{-\frac{x^{2}}{2}} dx\\=\int_{0,05}^0 \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2y^{2}}} -\frac{1}{y^2} dy \\=\int_0^{0,05} \frac{1}{\sqrt{2\pi}} e^{-\frac{1}{2y^{2}}} \frac{1}{y^2} dy$
